---
title: "STAT101C FINAL PROJECT"
author: "Team M&M: Megan Mets and Mounika Narayanan"
date: "5/24/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# read in the data
library(readr)
train <- read_csv("~/Downloads/train.csv") # 3000   17
test <- read_csv("~/Downloads/test.csv") # 1400   16
```

```{r}
######################### CHANGE THE FATAL COLUMN ##########################
train[train$Fatal == "U",]$Fatal <- "N"
train[train$Fatal == "F",]$Fatal <- "Y"
train$Fatal <- as.factor(train$Fatal)

# this is the clean function
clean <- function(mydata) {
  ######################### REMOVE UNNECESSARY COLUMNS ##########################
  mydata <- mydata[,-which(names(mydata) %in% c("NumberOfSubjects", "Department", "FullNarrative"))]
  
  ######################### CHANGE THE NOTES COLUMN ##########################
  notes <- tolower(mydata$Notes)
  hit <- rep("NA", length(notes))
  hit[grepl("no hit", notes)] <- 0
  hit[grepl("^hit$", notes)] <- 1
  hit[!grepl("no hit", notes) & (grepl("shot", notes) & grepl("hit", notes))] <- 1
  
  fatal <- rep("NA", length(notes))
  fatal[grepl("non-fatal", notes)] <- 0
  fatal[!grepl("non-fatal", notes) & grepl("fatal", notes)] <- 1
  fatal[grepl("killed", notes)] <- 1
  fatal[grepl("death", notes) & !grepl("no hit", notes)] <- 1
  
  weapon <- rep("NA", length(notes))
  weapon[grepl("toy gun|blank gun|unloaded gun|replica firearm", notes)] <- 0
  weapon[grepl("weapon|armed|firearm|handgun|knife|gun|revolver|rifle|knives|shotgun|grenade|bb gun|replica gun", notes) 
         & !grepl("officer's", notes)] <- 1
  
  ######################### CHANGE THE OFFICERGENDER COLUMN ##########################
  library(dplyr)
  mydata$OfficerGender[is.na(mydata$OfficerGender)] <- "U"
  mydata$OfficerGender <- unlist(lapply(strsplit(mydata$OfficerGender, ";"), FUN = function(x) {
    names(which.max(table(x)))
  }))
  mydata$OfficerGender[mydata$OfficerGender == "Unknown"] <- "U"
  mydata$OfficerGender[mydata$OfficerGender == "MALE"] <- "M"
  mydata$OfficerGender[mydata$OfficerGender == "FEMALE"] <- "F"
  
  ######################### CHANGE THE OFFICERRACE COLUMN ##########################
  mydata$OfficerRace[is.na(mydata$OfficerRace)] <- "U"
  mydata$OfficerRace <- unlist(lapply(strsplit(mydata$OfficerRace, ";"), FUN = function(x) {
    names(which.max(table(x)))
  }))
  mydata$OfficerRace[mydata$OfficerRace == "Unknown"] <- "U"
  mydata$OfficerRace[mydata$OfficerRace == "WHITE"] <- "W"
  mydata$OfficerRace[mydata$OfficerRace == "BLACK"] <- "B"
  
  mydata$hit <- hit
  mydata$newfatal <- fatal
  mydata$weapon <- weapon
  
  return(mydata)
}

train <- clean(train)
dim(train) # 3000   17
test <- clean(test)
dim(test)  # 1400   16
```

######################### CHANGE THE DATE COLUMN ##########################
```{r}
####### IN TRAIN DATASET #######
# read the date column as a date
train$Date <- as.Date(train$Date, "%Y-%m-%d")

# make a month column in the dataset
library(lubridate)
train$Month <- month(train$Date)
# make a Year column in the dataset
train$Year <- year(train$Date)

# make a fall, winter, spring, summer column
## fall= sept to nov (9-11)
## winter= dec to feb (12-2)
## spring= march to may (3-5)
## summer= june to august (6-8)
train$season <- c()
train$season[train$Month == 3 | train$Month == 4 | train$Month == 5] <- "spring"
train$season[train$Month == 6 | train$Month == 7 | train$Month == 8] <- "summer"
train$season[train$Month == 9 | train$Month == 10 | train$Month == 11] <- "fall"
train$season[train$Month == 12 | train$Month == 1 | train$Month == 2] <- "winter"

####### IN TEST DATASET #######
test$Date <- as.Date(test$Date, "%Y-%m-%d")

# make a month column in the dataset
library(lubridate)
test$Month <- month(test$Date)
test$Year <- year(test$Date)

# make a fall, winter, spring, summer column
## fall= sept to nov (9-11)
## winter= dec to feb (12-2)
## spring= march to may (3-5)
## summer= june to august (6-8)
test$season <- c()
test$season[test$Month == 3 | test$Month == 4 | test$Month == 5] <- "spring"
test$season[test$Month == 6 | test$Month == 7 | test$Month == 8] <- "summer"
test$season[test$Month == 9 | test$Month == 10 | test$Month == 11] <- "fall"
test$season[test$Month == 12 | test$Month == 1 | test$Month == 2] <- "winter"

# check new dimensions, added 2 columns: month and season
dim(train) # 3000   20
dim(test)  # 1400   19
```

######################### CHANGE THE SUBJECTAGE COLUMN ##########################
```{r}
####### IN TRAIN DATASET #######
# categorize into age ranges that makes sense
table(train$SubjectAge)
# 7 groups: 0-19, 20-29, 30-39, 40-49, 50-59, 60+, U
## making the 0-19 group
train$SubjectAge[train$SubjectAge == "Juvenile"] <- "0-19"
train$SubjectAge[train$SubjectAge == "4.0"] <- "0-19"
train$SubjectAge[train$SubjectAge == "12.0"] <- "0-19"
train$SubjectAge[train$SubjectAge == "13.0"] <- "0-19"
train$SubjectAge[train$SubjectAge == "14.0"] <- "0-19"
train$SubjectAge[train$SubjectAge == "15.0"] <- "0-19"
train$SubjectAge[train$SubjectAge == "16.0"] <- "0-19"
train$SubjectAge[train$SubjectAge == "17.0"] <- "0-19"
train$SubjectAge[train$SubjectAge == "18.0"] <- "0-19"
train$SubjectAge[train$SubjectAge == "19.0"] <- "0-19"
## making the 20-29 group
train$SubjectAge[train$SubjectAge == "20.0"] <- "20-29"
train$SubjectAge[train$SubjectAge == "21.0"] <- "20-29"
train$SubjectAge[train$SubjectAge == "22.0"] <- "20-29"
train$SubjectAge[train$SubjectAge == "23.0"] <- "20-29"
train$SubjectAge[train$SubjectAge == "24.0"] <- "20-29"
train$SubjectAge[train$SubjectAge == "25.0"] <- "20-29"
train$SubjectAge[train$SubjectAge == "26.0"] <- "20-29"
train$SubjectAge[train$SubjectAge == "27.0"] <- "20-29"
train$SubjectAge[train$SubjectAge == "28.0"] <- "20-29"
train$SubjectAge[train$SubjectAge == "29.0"] <- "20-29"
train$SubjectAge[train$SubjectAge == "21-23"] <- "20-29"
## making the 30-39 group
train$SubjectAge[train$SubjectAge == "30.0"] <- "30-39"
train$SubjectAge[train$SubjectAge == "31.0"] <- "30-39"
train$SubjectAge[train$SubjectAge == "32.0"] <- "30-39"
train$SubjectAge[train$SubjectAge == "33.0"] <- "30-39"
train$SubjectAge[train$SubjectAge == "34.0"] <- "30-39"
train$SubjectAge[train$SubjectAge == "35.0"] <- "30-39"
train$SubjectAge[train$SubjectAge == "36.0"] <- "30-39"
train$SubjectAge[train$SubjectAge == "37.0"] <- "30-39"
train$SubjectAge[train$SubjectAge == "38.0"] <- "30-39"
train$SubjectAge[train$SubjectAge == "39.0"] <- "30-39"
## making the 40-49 group
train$SubjectAge[train$SubjectAge == "40.0"] <- "40-49"
train$SubjectAge[train$SubjectAge == "41.0"] <- "40-49"
train$SubjectAge[train$SubjectAge == "42.0"] <- "40-49"
train$SubjectAge[train$SubjectAge == "43.0"] <- "40-49"
train$SubjectAge[train$SubjectAge == "44.0"] <- "40-49"
train$SubjectAge[train$SubjectAge == "45.0"] <- "40-49"
train$SubjectAge[train$SubjectAge == "46.0"] <- "40-49"
train$SubjectAge[train$SubjectAge == "47.0"] <- "40-49"
train$SubjectAge[train$SubjectAge == "48.0"] <- "40-49"
train$SubjectAge[train$SubjectAge == "49.0"] <- "40-49"
## making the 50-59 group
train$SubjectAge[train$SubjectAge == "50.0"] <- "50-59"
train$SubjectAge[train$SubjectAge == "51.0"] <- "50-59"
train$SubjectAge[train$SubjectAge == "52.0"] <- "50-59"
train$SubjectAge[train$SubjectAge == "53.0"] <- "50-59"
train$SubjectAge[train$SubjectAge == "54.0"] <- "50-59"
train$SubjectAge[train$SubjectAge == "55.0"] <- "50-59"
train$SubjectAge[train$SubjectAge == "56.0"] <- "50-59"
train$SubjectAge[train$SubjectAge == "57.0"] <- "50-59"
train$SubjectAge[train$SubjectAge == "58.0"] <- "50-59"
train$SubjectAge[train$SubjectAge == "59.0"] <- "50-59"
## making the 60+ group
train$SubjectAge[train$SubjectAge == "60.0"] <- "60+"
train$SubjectAge[train$SubjectAge == "61.0"] <- "60+"
train$SubjectAge[train$SubjectAge == "62.0"] <- "60+"
train$SubjectAge[train$SubjectAge == "63.0"] <- "60+"
train$SubjectAge[train$SubjectAge == "64.0"] <- "60+"
train$SubjectAge[train$SubjectAge == "65.0"] <- "60+"
train$SubjectAge[train$SubjectAge == "66.0"] <- "60+"
train$SubjectAge[train$SubjectAge == "67.0"] <- "60+"
train$SubjectAge[train$SubjectAge == "68.0"] <- "60+"
train$SubjectAge[train$SubjectAge == "69.0"] <- "60+"
train$SubjectAge[train$SubjectAge == "60-69"] <- "60+"
train$SubjectAge[train$SubjectAge == "73.0"] <- "60+"
## N/A, U, and UNKNOWN go into category U (i.e. unknown)
train$SubjectAge[train$SubjectAge == "N/A"] <- "U"
train$SubjectAge[train$SubjectAge == "UNKNOWN"] <- "U"


####### IN TEST DATASET #######
table(test$SubjectAge)
# 7 groups: 0-19, 20-29, 30-39, 40-49, 50-59, 60+, U
## making the 0-19 group
test$SubjectAge[test$SubjectAge == "13.0"] <- "0-19"
test$SubjectAge[test$SubjectAge == "14.0"] <- "0-19"
test$SubjectAge[test$SubjectAge == "15.0"] <- "0-19"
test$SubjectAge[test$SubjectAge == "16.0"] <- "0-19"
test$SubjectAge[test$SubjectAge == "17.0"] <- "0-19"
test$SubjectAge[test$SubjectAge == "18.0"] <- "0-19"
test$SubjectAge[test$SubjectAge == "19.0"] <- "0-19"
## making the 20-29 group
test$SubjectAge[test$SubjectAge == "20.0"] <- "20-29"
test$SubjectAge[test$SubjectAge == "21.0"] <- "20-29"
test$SubjectAge[test$SubjectAge == "22.0"] <- "20-29"
test$SubjectAge[test$SubjectAge == "23.0"] <- "20-29"
test$SubjectAge[test$SubjectAge == "24.0"] <- "20-29"
test$SubjectAge[test$SubjectAge == "25.0"] <- "20-29"
test$SubjectAge[test$SubjectAge == "26.0"] <- "20-29"
test$SubjectAge[test$SubjectAge == "27.0"] <- "20-29"
test$SubjectAge[test$SubjectAge == "28.0"] <- "20-29"
test$SubjectAge[test$SubjectAge == "29.0"] <- "20-29"
## making the 30-39 group
test$SubjectAge[test$SubjectAge == "30.0"] <- "30-39"
test$SubjectAge[test$SubjectAge == "31.0"] <- "30-39"
test$SubjectAge[test$SubjectAge == "32.0"] <- "30-39"
test$SubjectAge[test$SubjectAge == "33.0"] <- "30-39"
test$SubjectAge[test$SubjectAge == "34.0"] <- "30-39"
test$SubjectAge[test$SubjectAge == "35.0"] <- "30-39"
test$SubjectAge[test$SubjectAge == "36.0"] <- "30-39"
test$SubjectAge[test$SubjectAge == "37.0"] <- "30-39"
test$SubjectAge[test$SubjectAge == "38.0"] <- "30-39"
test$SubjectAge[test$SubjectAge == "39.0"] <- "30-39"
## making the 40-49 group
test$SubjectAge[test$SubjectAge == "40.0"] <- "40-49"
test$SubjectAge[test$SubjectAge == "41.0"] <- "40-49"
test$SubjectAge[test$SubjectAge == "42.0"] <- "40-49"
test$SubjectAge[test$SubjectAge == "43.0"] <- "40-49"
test$SubjectAge[test$SubjectAge == "44.0"] <- "40-49"
test$SubjectAge[test$SubjectAge == "45.0"] <- "40-49"
test$SubjectAge[test$SubjectAge == "46.0"] <- "40-49"
test$SubjectAge[test$SubjectAge == "47.0"] <- "40-49"
test$SubjectAge[test$SubjectAge == "48.0"] <- "40-49"
test$SubjectAge[test$SubjectAge == "49.0"] <- "40-49"
## making the 50-59 group
test$SubjectAge[test$SubjectAge == "50.0"] <- "50-59"
test$SubjectAge[test$SubjectAge == "51.0"] <- "50-59"
test$SubjectAge[test$SubjectAge == "52.0"] <- "50-59"
test$SubjectAge[test$SubjectAge == "53.0"] <- "50-59"
test$SubjectAge[test$SubjectAge == "54.0"] <- "50-59"
test$SubjectAge[test$SubjectAge == "55.0"] <- "50-59"
test$SubjectAge[test$SubjectAge == "56.0"] <- "50-59"
test$SubjectAge[test$SubjectAge == "57.0"] <- "50-59"
test$SubjectAge[test$SubjectAge == "58.0"] <- "50-59"
test$SubjectAge[test$SubjectAge == "59.0"] <- "50-59"
## making the 60+ group
test$SubjectAge[test$SubjectAge == "61.0"] <- "60+"
test$SubjectAge[test$SubjectAge == "62.0"] <- "60+"
test$SubjectAge[test$SubjectAge == "63.0"] <- "60+"
test$SubjectAge[test$SubjectAge == "64.0"] <- "60+"
test$SubjectAge[test$SubjectAge == "68.0"] <- "60+"
test$SubjectAge[test$SubjectAge == "70.0"] <- "60+"
test$SubjectAge[test$SubjectAge == "71.0"] <- "60+"
test$SubjectAge[test$SubjectAge == "72.0"] <- "60+"
test$SubjectAge[test$SubjectAge == "78.0"] <- "60+"
test$SubjectAge[test$SubjectAge == "86.0"] <- "60+"
## N/A, U, and UNKNOWN go into category U (i.e. unknown)
test$SubjectAge[test$SubjectAge == "N/A"] <- "U"
test$SubjectAge[test$SubjectAge == "UNKNOWN"] <- "U"

# check new tables, there are enough values for each category
table(train$SubjectAge)
table(test$SubjectAge)
```

######################### CHANGE THE SUBJECTARMED COLUMN ##########################
```{r}
# need to change the NA's to U
####### IN TRAIN DATASET #######
unique(train$SubjectArmed)
table(train$SubjectArmed)
sum(is.na(train$SubjectArmed)) # 322
train$SubjectArmed[is.na(train$SubjectArmed)] <- "U"

####### IN TEST DATASET #######
unique(test$SubjectArmed)
table(test$SubjectArmed)
sum(is.na(test$SubjectArmed))
test$SubjectArmed[is.na(test$SubjectArmed)] <- "U"
```

######################### CHANGE THE SUBJECTRACE COLUMN ##########################
```{r}
# need to change the NA's to U
####### IN TRAIN DATASET #######
unique(train$SubjectRace)
table(train$SubjectRace)
sum(is.na(train$SubjectRace)) # 1
train$SubjectRace[is.na(train$SubjectRace)] <- "U"

####### IN TEST DATASET #######
unique(test$SubjectRace)
table(test$SubjectRace)
sum(is.na(test$SubjectRace)) # 0 dont need to change anything
```

######################### CHANGE THE SUBJECTGENDER COLUMN ##########################
```{r}
# need to change the NA's to U
####### IN TRAIN DATASET #######
unique(train$SubjectGender)
table(train$SubjectGender)
train$SubjectGender[train$SubjectGender == "N/A"] <- "U"

####### IN TEST DATASET #######
unique(test$SubjectGender)
table(test$SubjectGender)
sum(is.na(test$SubjectGender)) # 0, no NA's to change
# change the 1 M;U to M
test$SubjectGender[test$SubjectGender == "M;U"] <- "M"
```

######################### REMOVE THE NOTES COLUMN ##########################
```{r}
train_clean <- train
train_clean <- train_clean[,-14]
test_clean <- test
test_clean <- test_clean[,-13]
dim(train_clean)  # 3000   19
dim(test_clean)  # 1400   18
```

######################### RECODING OFFICERRACE AND OFFICERGENDER ##########################
```{r}
test_clean$OfficerGender[test_clean$OfficerGender == "Male"] <- "M"
train_clean$OfficerGender[train_clean$OfficerGender == "M:M"] <- "M"
train_clean$OfficerGender[train_clean$OfficerGender == "N"] <- "U"

train_clean$OfficerRace[train_clean$OfficerRace == "A/PI"] <- "A"
train_clean$OfficerRace[train_clean$OfficerRace == "W/ H"] <- "W"
train_clean$OfficerRace[train_clean$OfficerRace == "A/W"] <- "A"
train_clean$OfficerRace[train_clean$OfficerRace == "AI/AN"] <- "A"
train_clean$OfficerRace[train_clean$OfficerRace == "ASIAN"] <- "A"
train_clean$OfficerRace[train_clean$OfficerRace == "H/L"] <- "L"
train_clean$OfficerRace[train_clean$OfficerRace == "NA/W"] <- "W"
train_clean$OfficerRace[train_clean$OfficerRace == "O"] <- "U"

test_clean$OfficerRace[test_clean$OfficerRace == "A/PI"] <- "A"
test_clean$OfficerRace[test_clean$OfficerRace == "Other:B"] <- "B"
test_clean$OfficerRace[test_clean$OfficerRace == "m/m"] <- "U"
test_clean$OfficerRace[test_clean$OfficerRace == "A/W"] <- "A"
test_clean$OfficerRace[test_clean$OfficerRace == "A/PI Unknown"] <- "A"
test_clean$OfficerRace[test_clean$OfficerRace == "AI/AN"] <- "A"
test_clean$OfficerRace[test_clean$OfficerRace == "M"] <- "U"
test_clean$OfficerRace[test_clean$OfficerRace == "O"] <- "U"
```

######################### CHANGE THE NUMBEROFOFFICERS COLUMN ##########################
```{r}
# change to 2 levels: 1-3 and 4+
train_clean$NumberOfOfficers[train_clean$NumberOfOfficers == 0] <- 1
hist(train_clean$NumberOfOfficers)
train_clean$fatfac <- ifelse(train_clean$Fatal == "Y", 1, 0)
table(train_clean$fatfac, train_clean$NumberOfOfficers)
train_clean$NumberOfOfficers[is.na(train_clean$NumberOfOfficers)] <- "U"
test_clean$NumberOfOfficers[is.na(test_clean$NumberOfOfficers)] <- "U"

train_clean$NumberOfOfficers[train_clean$NumberOfOfficers == 1] <- "1-3"
train_clean$NumberOfOfficers[train_clean$NumberOfOfficers == 2] <- "1-3"
train_clean$NumberOfOfficers[train_clean$NumberOfOfficers == 3] <- "1-3"
train_clean$NumberOfOfficers[train_clean$NumberOfOfficers == 4] <- "4+"
train_clean$NumberOfOfficers[train_clean$NumberOfOfficers == 5] <- "4+"
train_clean$NumberOfOfficers[train_clean$NumberOfOfficers == 6] <- "4+"
train_clean$NumberOfOfficers[train_clean$NumberOfOfficers == 7] <- "4+"
train_clean$NumberOfOfficers[train_clean$NumberOfOfficers == 8] <- "4+"
train_clean$NumberOfOfficers[train_clean$NumberOfOfficers == 9] <- "4+"
train_clean$NumberOfOfficers[train_clean$NumberOfOfficers == 10] <- "4+"
train_clean$NumberOfOfficers[train_clean$NumberOfOfficers == 11] <- "4+"
train_clean$NumberOfOfficers[train_clean$NumberOfOfficers == 13] <- "4+"
train_clean$NumberOfOfficers[train_clean$NumberOfOfficers == 15] <- "4+"
train_clean$NumberOfOfficers <- as.factor(train_clean$NumberOfOfficers)

test_clean$NumberOfOfficers[test_clean$NumberOfOfficers == 1] <- "1-3"
test_clean$NumberOfOfficers[test_clean$NumberOfOfficers == 2] <- "1-3"
test_clean$NumberOfOfficers[test_clean$NumberOfOfficers == 3] <- "1-3"
test_clean$NumberOfOfficers[test_clean$NumberOfOfficers == 4] <- "4+"
test_clean$NumberOfOfficers[test_clean$NumberOfOfficers == 5] <- "4+"
test_clean$NumberOfOfficers[test_clean$NumberOfOfficers == 6] <- "4+"
test_clean$NumberOfOfficers[test_clean$NumberOfOfficers == 7] <- "4+"
test_clean$NumberOfOfficers[test_clean$NumberOfOfficers == 8] <- "4+"
test_clean$NumberOfOfficers[test_clean$NumberOfOfficers == 9] <- "4+"
test_clean$NumberOfOfficers[test_clean$NumberOfOfficers == 10] <- "4+"
test_clean$NumberOfOfficers[test_clean$NumberOfOfficers == 11] <- "4+"
test_clean$NumberOfOfficers[test_clean$NumberOfOfficers == 12] <- "4+"
test_clean$NumberOfOfficers[test_clean$NumberOfOfficers == 16] <- "4+"
test_clean$NumberOfOfficers[test_clean$NumberOfOfficers == 23] <- "4+"
test_clean$NumberOfOfficers <- as.factor(test_clean$NumberOfOfficers)

train_clean$NumberOfOfficers <- as.factor(train_clean$NumberOfOfficers)
unique(train_clean$NumberOfOfficers)
unique(test_clean$NumberOfOfficers)
```

######################### CHANGING COLUMN TYPES ##########################
```{r}
test_clean$SubjectArmed <- as.factor(test_clean$SubjectArmed)
test_clean$SubjectRace <- as.factor(test_clean$SubjectRace)
test_clean$SubjectGender <- as.factor(test_clean$SubjectGender)
test_clean$SubjectAge <- as.factor(test_clean$SubjectAge)
test_clean$OfficerRace <- as.factor(test_clean$OfficerRace)
test_clean$OfficerGender <- as.factor(test_clean$OfficerGender)
test_clean$hit <- as.factor(test_clean$hit)
test_clean$newfatal <- as.factor(test_clean$newfatal)
test_clean$weapon <- as.factor(test_clean$weapon)
test_clean$Month <- as.factor(test_clean$Month)
test_clean$season <- as.factor(test_clean$season)
test_clean$City <- as.factor(test_clean$City)

train_clean$SubjectArmed <- as.factor(train_clean$SubjectArmed)
train_clean$SubjectRace <- as.factor(train_clean$SubjectRace)
train_clean$SubjectGender <- as.factor(train_clean$SubjectGender)
train_clean$SubjectAge <- as.factor(train_clean$SubjectAge)
train_clean$OfficerRace <- as.factor(train_clean$OfficerRace)
train_clean$OfficerGender <- as.factor(train_clean$OfficerGender)
train_clean$hit <- as.factor(train_clean$hit)
train_clean$newfatal <- as.factor(train_clean$newfatal)
train_clean$weapon <- as.factor(train_clean$weapon)
train_clean$season <- as.factor(train_clean$season)
train_clean$Month <- as.factor(train_clean$Month)
train_clean$City <- as.factor(train_clean$City)
```

######################### CHANGE THE NUMBEROFSHOTS COLUMN ##########################
```{r}
## remove this column because 72% are NA
train_clean <- train_clean[,-9]
test_clean <- test_clean[,-8]
```

######################### CHANGE THE NATUREOFSTOP COLUMN ##########################
```{r}
sum(is.na(train_clean$NatureOfStop))
1929/3000  ## 64% NA

# remove column, no benefit
train_clean <- train_clean[,-8]
train_clean <- train_clean[,-18]
test_clean <- test_clean[,-7]

dim(train_clean)  # 3000   17
dim(test_clean) # 1400   16
```

######################### FIXING MORE COLUMNS WITH NAs ##########################

```{r}
train_clean$weapon <- as.character(train_clean$weapon)
train_clean$weapon[is.na(train_clean$weapon)] <- "U"
train_clean$NumberOfOfficers <- as.factor(train_clean$NumberOfOfficers)

train_clean$Month <- as.character(train_clean$Month)
train_clean$Month[is.na(train_clean$Month)] <- "U"
train_clean$Month <- as.factor(train_clean$Month)

train_clean$season <- as.character(train_clean$season)
train_clean$season[is.na(train_clean$season)] <- "U"
train_clean$season <- as.factor(train_clean$season)
str(train_clean$season)

train_clean$Date <- as.character(train_clean$Date)
train_clean$Date[is.na(train_clean$Date)] <- "U"
train_clean$Date <- as.factor(train_clean$Date)
str(train_clean$Date)

test_clean$weapon <- as.character(test_clean$weapon)
test_clean$weapon[test_clean$weapon == "NA"] <- "U"
test_clean$NumberOfOfficers <- as.factor(test_clean$NumberOfOfficers)

test_clean$Month <- as.character(test_clean$Month)
test_clean$Month[is.na(test_clean$Month)] <- "U"
test_clean$Month <- as.factor(test_clean$Month)

test_clean$season <- as.character(test_clean$season)
test_clean$season[is.na(test_clean$season)] <- "U"
test_clean$season <- as.factor(test_clean$season)
str(test_clean$season)

test_clean$Date <- as.character(test_clean$Date)
test_clean$Date[is.na(test_clean$Date)] <- "U"
test_clean$Date <- as.factor(test_clean$Date)
str(test_clean$Date)

train_clean$hit <- as.character(train_clean$hit)
train_clean$hit[train_clean$hit == "NA"] <- "U"
train_clean$hit <- as.factor(train_clean$hit)

train_clean$weapon <- as.character(train_clean$weapon)
train_clean$weapon[train_clean$weapon == "NA"] <- "U"
train_clean$weapon <- as.factor(train_clean$weapon)

train_clean$newfatal <- as.character(train_clean$newfatal)
train_clean$newfatal[train_clean$newfatal == "NA"] <- "U"
train_clean$newfatal <- as.factor(train_clean$newfatal)

train_clean$Year <- as.character(train_clean$Year)
train_clean$Year[is.na(train_clean$Year)] <- "U"
train_clean$Year <- as.factor(train_clean$Year)

test_clean$hit <- as.character(test_clean$hit)
test_clean$hit[test_clean$hit == "NA"] <- "U"
test_clean$hit <- as.factor(test_clean$hit)

test_clean$newfatal <- as.character(test_clean$newfatal)
test_clean$newfatal[test_clean$newfatal == "NA"] <- "U"
test_clean$newfatal <- as.factor(test_clean$newfatal)

test_clean$weapon <- as.factor(test_clean$weapon)

test_clean$Year[is.na(test_clean$Year)] <- "U"
test_clean$Year <- as.factor(test_clean$Year)

str(train_clean)
str(test_clean)
dim(train_clean) # 3000   17
dim(test_clean)  # 1400   16

train_clean$Month <- as.character(train_clean$Month)
train_clean$Month[train_clean$Month == "U"] <- NA
train_clean$Month <- as.numeric(train_clean$Month)

test_clean$Month <- as.character(test_clean$Month)
test_clean$Month[test_clean$Month == "U"] <- NA
test_clean$Month <- as.numeric(test_clean$Month)

str(train_clean)
str(test_clean)
```

######################### BEST MODEL ##########################
```{r}
# BOOSTING
train_clean$Fatal <- as.numeric(train_clean$Fatal)
train_clean$Fatal[train_clean$Fatal == 1] <- 0
train_clean$Fatal[train_clean$Fatal == 2] <- 1

junk1 <- train_clean
junk2<- test_clean

junk1$Year <- as.character(junk1$Year)
junk1$Year[junk1$Year == "U"] <- NA
junk1$Year <- as.numeric(junk1$Year)

junk2$Year <- as.character(junk2$Year)
junk2$Year[junk2$Year == "U"] <- NA
junk2$Year <- as.numeric(junk2$Year)

set.seed(123456)
library(gbm)
boost.t <- gbm(Fatal ~ SubjectArmed + SubjectRace + SubjectGender + SubjectAge + 
                      OfficerRace + OfficerGender + Year + hit + newfatal + weapon +
                      NumberOfOfficers, data= junk1, distribution="gaussian", n.trees =5000 , interaction.depth =4)
summary(boost.t)

yhat.test <- predict(boost.t, newdata= junk2, n.trees=5000)
boost <- round(yhat.test)

boost[boost == 0] <- "No"
boost[boost == "1"] <- "Yes"
boost[junk2$hit == "0"] <- "No"
boost[junk2$newfatal == "1"] <- "Yes"
boost[junk2$newfatal == "0"] <- "No"

predz2 <- cbind(junk2$id, boost)
predz2 <- as.data.frame(predz2)
names(predz2) <- c("id", "Fatal")

table(predz2$Fatal)

write.csv(predz2, "june8_best2.csv", row.names = F)
```

